Reduced plasticity and opportunity for selection in terrestrial ectotherm populations under climate change

### 2023-05-14 ms.qmd
re-rendered MS with updated climate results

 col <- c("#0871B9", "#F3B40D", "#1BB908")
p8.1 <- bubble_plot(model3.9, mod = "cv_c", group = "record_num", by = "habitat", data = lnRRQ10_data_wild_acclim, ylab = TeX("$lnRR_{Q_{10}}$"), xlab = "Thermal Coefficient of Variation (CV)", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text, aes(x = x, y = y-0.3, label = label2), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y-0.6, label = p), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y+4, label = title), size = 5) + 
theme(legend.position = "none", strip.background = element_blank(), strip.text = element_blank(),axis.text = element_text(size = size_axis), axis.title = element_text(size = size_axis))

@@

 col <- c("#0871B9", "#F3B40D", "#1BB908")
p9.1 <- bubble_plot(model4.9, mod = "cv_c", group = "record_num", by = "habitat", data = lnCVRQ10_data_wild_acclim, ylab = TeX("$lnCVR_{Q_{10}}$"), xlab = "Thermal Coefficient of Variation (CV)", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text, aes(x = x, y = y-1, label = label2), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y-2, label = p), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y+10, label = title, size = 5)) + 
theme(legend.position = "none", strip.background = element_blank(), strip.text = element_blank(),axis.text = element_text(size = size_axis), axis.title = element_text(size = size_axis))

@@

p9.2 <- bubble_plot(model4.8, mod = "acf_all_c", group = "record_num", by = "habitat", data = lnCVRQ10_data_wild_acclim, ylab = TeX("$lnCVR_{Q_{10}}$"), xlab = "Thermal Predictability", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text2, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text2, aes(x = x, y = y-0.6, label = label2), size = 5) +
geom_text(data = dat_text2, aes(x = x, y = y-1.2, label = p), size = 5) +
theme(legend.position = "bottom", 
      strip.background = element_blank(), 
      strip.text = element_blank(), 
      axis.text = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      legend.text = element_text(size = 16),
      legend.title = element_text(size = 16))

@@

   model5_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  c_accl_time + type*habitat + t2(long.x, lat.y) + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_cvr)) + (1 | spp) + (1|trait) + (1|obs), data2 = list(A_wild_cvr=A_wild_cvr), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.99, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),])

@@

# Fitted by Lat/long
  ggplot(data = data_w_pred, aes(y = Estimate, x = lat.y)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Predicted Effect Size") + theme_classic()

  ggplot(data = data_w_pred, aes(y = Estimate, x = long.x)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Predicted Effect Size") + theme_classic()

# Observed by Lat/long
  ggplot(data = data_w_pred, aes(y = effect_size, x = lat.y)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Observed Effect Size") + theme_classic()

  ggplot(data = data_w_pred, aes(y = effect_size, x = long.x)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Observed Effect Size") + theme_classic()

@@

  terrest_data <- sub_data  %>% filter(habitat == "t")  %>% distinct(spp, species_full, habitat, type, lat.y, long.x)
  new_data_t <- data.frame(
      c_accl_time = 0,
      V = 0,
      trait = 0,
      obs = 0, 
      record_num = 0
    )

@@

#############################
## Marine ectotherms  
#############################
  marine_data <- sub_data  %>% filter(habitat == "m") %>% distinct(spp, species_full, habitat, type, lat.y, long.x)

@@

#############################
## Freshwater ectotherms
#############################
  fresh_data <- sub_data  %>% filter(habitat == "f")%>% distinct(spp, species_full, habitat, type, lat.y, long.x)

@@

map1_current <- ggplot() +
  geom_map(
    data = world_map, map = world_map,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray") +
  expand_limits(x = world_map$long, y = world_map$lat) +
  coord_map("moll") +
  geom_point(
    data = all_pred  %>% filter(type == "acclim"),
    aes(long.x, lat.y, color = Estimate),
    alpha = 0.7, size = 5
  ) + labs(x = "Longitude", y = "Latitude", color = TeX("$lnCVR_{Q_{10}}$")) + scale_color_viridis() + theme_map(base_size = 22)

@@

#############
  ## lnVRQ10
#############
##################################################
  # Organise the data and prune tree
#################################################
  lnVRQ10_data <- data_long2 %>% filter(grepl("lnVR", name) & !is.na(effect_size) & !is.na(V) & !V == 0) %>% 
                              mutate(spp = species_full, c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE))

>>>>

 col <- c("#0871B9", "#F3B40D", "#1BB908")
p8.1 <- bubble_plot(model3.9, mod = "cv_c", group = "record_num", by = "habitat", ylab = TeX("$lnRR_{Q_{10}}$"), xlab = "Thermal Coefficient of Variation (CV)", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text, aes(x = x, y = y-0.3, label = label2), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y-0.6, label = p), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y+4, label = title), size = 5) + 
theme(legend.position = "none", strip.background = element_blank(), strip.text = element_blank(),axis.text = element_text(size = size_axis), axis.title = element_text(size = size_axis))

@@

 col <- c("#0871B9", "#F3B40D", "#1BB908")
p9.1 <- bubble_plot(model4.9, mod = "cv_c", group = "record_num", by = "habitat",  ylab = TeX("$lnCVR_{Q_{10}}$"), xlab = "Thermal Coefficient of Variation (CV)", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text, aes(x = x, y = y-1, label = label2), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y-2, label = p), size = 5) +
geom_text(data = dat_text, aes(x = x, y = y+10, label = title, size = 5)) + 
theme(legend.position = "none", strip.background = element_blank(), strip.text = element_blank(),axis.text = element_text(size = size_axis), axis.title = element_text(size = size_axis))

@@

p9.2 <- bubble_plot(model4.8, mod = "acf_all_c", group = "record_num", by = "habitat",  ylab = TeX("$lnCVR_{Q_{10}}$"), xlab = "Thermal Predictability", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text2, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text2, aes(x = x, y = y-0.6, label = label2), size = 5) +
geom_text(data = dat_text2, aes(x = x, y = y-1.2, label = p), size = 5) +
theme(legend.position = "bottom", 
      strip.background = element_blank(), 
      strip.text = element_blank(), 
      axis.text = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      legend.text = element_text(size = 16),
      legend.title = element_text(size = 16))

@@

   model5_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  c_accl_time + type*habitat + t2(Longitude, Latitude) + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_cvr)) + (1 | spp) + (1|trait) + (1|obs), data2 = list(A_wild_cvr=A_wild_cvr), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.99, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),])

@@

# Fitted by Lat/long Longitude, Latitude
  ggplot(data = data_w_pred, aes(y = Estimate, x = Latitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Predicted Effect Size") + theme_classic()

  ggplot(data = data_w_pred, aes(y = Estimate, x = Longitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Predicted Effect Size") + theme_classic()

# Observed by Lat/long
  ggplot(data = data_w_pred, aes(y = effect_size, x = Latitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Observed Effect Size") + theme_classic()

  ggplot(data = data_w_pred, aes(y = effect_size, x = Longitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Observed Effect Size") + theme_classic()

@@

  terrest_data <- sub_data  %>% filter(habitat == "t")  %>% distinct(spp, species_full, habitat, type, Latitude, Longitude)
  new_data_t <- data.frame(
      c_accl_time = 0,
      V = 0,
      trait = 0,
      obs = 0, 
      record_num = 0
    )

@@

#############################
## Marine ectotherms  
#############################
  marine_data <- sub_data  %>% filter(habitat == "m") %>% distinct(spp, species_full, habitat, type, Latitude, Longitude)

@@

#############################
## Freshwater ectotherms
#############################
  fresh_data <- sub_data  %>% filter(habitat == "f")%>% distinct(spp, species_full, habitat, type, Latitude, Longitude)

@@

map1_current <- ggplot() +
  geom_map(
    data = world_map, map = world_map,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray") +
  expand_limits(x = world_map$long, y = world_map$lat) +
  coord_map("moll") +
  geom_point(
    data = all_pred  %>% filter(type == "acclim"),
    aes(Longitude, Latitude, color = Estimate),
    alpha = 0.7, size = 5
  ) + labs(x = "Longitude", y = "Latitude", color = TeX("$lnCVR_{Q_{10}}$")) + scale_color_viridis() + theme_map(base_size = 22)

@@

#############
  ## lnVRQ10
#############
##################################################
  # Organise the data and prune tree
#################################################
  lnVRQ10_data <- data_long2 %>% filter(grepl("lnVR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period)) %>% 
                              mutate(spp = species_full, c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE))
